\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{float}
\usepackage{tikz}
\usepackage{mlmodern}
\usepackage{subfig}
\usepackage{pgfornament}
\usepackage{graphicx}
\usepackage[some]{background}
\usepackage[T1]{fontenc}
\usepackage[left=2.5cm,right=2.5cm]{geometry}
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{hyperref, xurl}
\usepackage{qrcode}
\usepackage{authblk} % for author info
\usepackage{svg}
\usepackage{listings}
\usepackage{epigraph} % used for the quotes

\usepackage[sorting=none]{biblatex}
\addbibresource{references.bib}

\newcommand{\maniflow}{\texttt{maniflow}}
\newcommand{\code}[1]{\texttt{#1}} % inline code setting
\renewcommand{\lstlistingname}{Code}

\definecolor{codebg}{RGB}{230,230,230}
\lstset {
    breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
    breaklines=true,                 % sets automatic line breaking
    captionpos=b,                    % sets the caption-position to bottom
    tabsize=4,
    %showstringspaces=false,
    numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
    numbersep=2pt,                   % how far the line-numbers are from the code
    numberstyle=\tiny, % the style that is used for the line-numbers
    %upquote=true,
    backgroundcolor = \color{codebg},
    basicstyle=\small\ttfamily, % basic font setting
    emph={int,char,double,float,unsigned,void,bool},
    %emphstyle={\color{blue}},
    escapechar=\&,
    columns=fixed
}

\usepackage{collectbox}

\makeatletter
\newcommand{\boxedcode}{%
    \collectbox{%
        \setlength{\fboxsep}{5pt}%
        \fbox{\BOXCONTENT}%
    }%
}
\makeatother

\newtheorem{defi}{Definition}
\newtheorem{ex}{Example}
\newtheorem{prop}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{kor}{Corollary}

\newcommand{\N}{\mathbb{N}}

\newcommand{\R}{\mathbb{R}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\E}{\mathcal{E}}

\author[]{Felix Widmaier}
\author[]{Robin Mark Riegraf}
\author[]{Yangshan Xiang}
\author[]{Minming Zhao}
\affil[]{Georg-August-Universität Göttingen}
%\date{}

\usepackage{lipsum}

\begin{document}
%\input{frontpage}
%\input{sections/frontmatter}
\newpage
%\input{sections/libinfo}
%\input{sections/introduction}
\section{Introduction}
Mesh simplification is a fundamental technique in computer graphics and computational geometry, employed to optimize the performance of various applications, such as 3D modeling, simulation, and rendering. By reducing the complexity of a mesh while preserving its essential features, mesh simplification achieves a balance between computational efficiency and visually staying true to the original.

The process of mesh simplification involves iteratively removing vertices, edges, or faces from the original mesh while minimizing the perceptible loss of detail. Various algorithms and methodologies have been developed to accomplish this task, ranging from straightforward decimation techniques to sophisticated error-driven approaches.

Furthermore, mesh simplification is not only beneficial for real-time rendering and interactive applications but also plays a crucial role in data storage, transmission, and manipulation. By generating simplified representations of complex meshes, the storage requirements are minimized, facilitating efficient data exchange and manipulation across different platforms and devices.

\section{Implemented Approach}
\subsection{Approach to Mesh Simplification}
In our approach we deploy vertex contraction as described in "Surface Simplification Using Quadric Error Metrics". This way we cut down on the amount of vertices and hence faces. Technically, also the amount of edges is reduced but in our project the edges are only considered via vertices and faces.\\
The procedure has several steps. First, we have to find elegible pairs of vertices for the contraction. There are in turn two kinds of pairs. Either they are part of the same face, that is that they are connected by an edge. The alternative criterion for a valid pair with regard to contraction is that the vertices of the pair are reasonably close towards eachother. That 
\subsection{Description of the Code}
Our implemented code aims to perform mesh simplification by intelligently reducing the complexity of input meshes while preserving their essential geometric characteristics.

\section{Comparion with Other Approaches}

\section{Outlook and Future Works}

%\input{sections/facegraph}
%\input{sections/render}
%\input{sections/geometry}
\newpage
\printbibliography
\end{document}